Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2025-01-20 article.tex
Updates

It is noted that the displacement is assumed to vary linearly within each cell; hence, the displacement gradient and stress are constant within each cell.
In the current work, a unique definition of stress $\bb{\sigma}_{f_i}$ at each cell face $f_i$ is given as a weighted-averaged of values in the two cells ($\bb{\sigma}_P $, $\bb{\sigma}_{N_{f_i}}$) straddling the face \citep{Jasak1996}:
%The stress $\bb{\sigma}_{f_i}$ at an internal face is calculated by linear interpolation from the adjacent cell centres \citep{Jasak1996}:
\begin{eqnarray} \label{eq:stressInterp}
	\bb{\sigma}_{f_i} &=& w_{f_i} \bb{\sigma}_P + (1 - w_{f_i}) \bb{\sigma}_{N_{f_i}}
\end{eqnarray}
%where $\bb{\sigma}_{f_i}$ is the stress at an interface face $f_i$, 
where the interpolation weight is defined as $w_{f_i} = (\bb{n}_{f_i} \cdot [\bb{x}_{N_{f_i}} - \bb{x}_{f_i}])/(\bb{n}_{f_i} \cdot [\bb{x}_{N_{f_i}} - \bb{x}_{P} ])$; however, achieving second-order accuracy of the displacement field is independent of the value of the weights, and, for example, $w_{f_i} = \nicefrac{1}{2}$ would also be sufficient.
Similarly, the stress $\bb{\sigma}_{s_i}$ at a symmetry boundary face is calculated as
\begin{eqnarray} \label{eq:symm}
	\bb{\sigma}_{s_i}
		&=& \frac{1}{2} \left (\bb{\sigma}_P + \bb{R}_{s_i} \cdot \bb{\sigma}_{P} \right) \notag \\
		&=& \left (\textbf{I} - \bb{n}_{s_i} \otimes \bb{n}_{s_i} \right) \cdot \bb{\sigma}_P
\end{eqnarray}
where $\bb{R}_{s_i} \cdot \bb{\sigma}_{P}$ represents the mirror reflection of $\bb{\sigma}_P$ across the symmetry boundary face $s_i$.
The reflection tensor is $\bb{R}_{s_i} = \textbf{I} - 2 \bb{n}_{s_i} \otimes \bb{n}_{s_i}$ \citep{Demirdzic2022}, with $\bb{n}_{s_i}$ indicating the unit normal of the symmetry boundary face $s_i$.
From Equation \ref{eq:symm}, it is clear that shear stresses are zero on a symmetry plane boundary face.
Note from Equation \ref{eq:divStressDiscret} that the stress on displacement boundary faces is assumed to be equal to the stress $\bb{\sigma}_P$ at the centroid of cell $P$.

@@

\hl{magSf in OF}

@@

first-order accurate gradients are sufficient to preserve second-order accuracy of the cell-centre displacements.

@@

%The predicted $\sigma_{xx}$ stress distribution for a hexahedral mesh with $512\,000$ cells is shown in Figure \ref{fig:mms_stress}(a).
%The corresponding cell-wise $\sigma_{xx}$ error distribution is shown in Figure \ref{fig:mms_stress}(b).
%The errors of greatest magnitude ($43$ kPa) occur at the boundaries, corresponding to where the local truncation error is higher. % due to the use of one-sided differencing.
%\begin{figure}[htbp]
%	\centering
%	\subfigure[Predicted $\sigma_{xx}$ stress distribution]
%	{
%		\label{fig:mms_sxx}
%   		\includegraphics[height=0.35\textwidth]{figures/mms_sxx} 
%   	}
%	\subfigure[Cell-wise $\sigma_{xx}$ error distribution]
%	{
%		\label{fig:mms_sxx_diff}
%   		\includegraphics[height=0.35\textwidth]{figures/mms_sxx_diff}  
%   	}
%	\caption{Manufactured solution cube case: the predicted $\sigma_{xx}$ stress distribution on a cut-plane for the regular hexahedral mesh with $512\,000$ cells (left).}
%	\label{fig:mms_stress}
%\end{figure}

%\paragraph{Case 2: Out-of-plane bending of an elliptic plate}
%This 3-D, static, linear elastic test case (Figure \ref{fig:elliptic_plate}) consists of a thick elliptic plate (0.6 m thick) with a centred elliptic hole, with the inner and outer ellipses given as
%\begin{eqnarray}
%	\left(\frac{x}{2}\right)^2 + \left(\frac{y}{1}\right)^2 = 1 & \text{inner ellipse} \\
%	\left(\frac{x}{3.25}\right)^2 + \left(\frac{y}{2.75}\right)^2 = 1 & \text{outer ellipse}
%\end{eqnarray}
%The case has been described by the National Agency for Finite Element Methods and Standards (NAFEMS) \cite{Hitchings1987}, and analysed using finite volume procedures by \citet{Demirdzic1997a} and \citet{Cardiff2016a}.
%Symmetry allows one-quarter of the geometry to be simulated.
%A constant pressure of 1 MPa is applied to the upper surface, and the outer surface is fully clamped.
%The mechanical properties are: $E = 210$ GPa, $\nu = 0.3$.
%Six successively refined hexahedral meshes are used, with cell counts of 45, 472 (Figure \ref{fig:elliptic_plate}), $4\,140$, $34\,968$, $287\,280$ and $2\,438\,242$.
%\begin{figure}[htbp]
%   \centering
%%\includegraphics[width=0.5\textwidth]{figures/elliptic_plate.pdf} ellipticPlate-geometry
%%	\subfigure[Mesh containing 472 hexahedral cells]
%%	{
%		\includegraphics[width=0.45\textwidth]{figures/elliptic_plate.pdf} 
%%	}
%   \caption{Elliptic Plate geometry and mesh containing 472 hexahedral cells}
%   \label{fig:elliptic_plate}
%\end{figure}
%
%The prediction for the equivalent (von Mises) stress in the domain are shown in Figure \ref{fig:elliptic_plate_sigmaEq}(a), and the values along the line $r = \sqrt{x^2 + y^2} = 2.1$ m, $z = 0.3$ m are shown in Figure \ref{fig:elliptic_plate_sigmaEq}(b).
%The results from the finest grid in \citet{Demirdzic1997a} are given for comparison, where good agreement is seen;
%the small offset between the finest mesh predictions and those from \citet{Demirdzic1997a} are likely due to errors introduced when extracting the \citet{Demirdzic1997a} results using WebPlotDigitizer \citep{WebPlotDigitizer}.
%The sample values have been calculated by generating 40 uniformly spaced angles between 0 and $\nicefrac{\pi}{2}$ along the line, finding the cell of each sampled point, and extrapolating from the cell-centred value using the field gradient.
%\begin{figure}[htbp]
%   \centering
%   %\includegraphics[width=0.5\textwidth]{figures/elliptic_plate_sigmaEq_line.pdf} 
%	\subfigure[Equivalent stress distribution for the mesh with $2\,438\,242$ cells. The deformation is scaled by $1\,000$ and a translucent line indicates the outline of the undeformed geometry]
%	{
%		\includegraphics[width=0.45\textwidth]{figures/elliptic_plate_sigmaEq}  
%	}
%	\subfigure[Equivalent stress along the line $r = 2.1$, $z = 0.3$ m]
%	{
%		\includegraphics[width=0.45\textwidth]{figures/elliptic_plate_sigmaEq_line.pdf} 
%	}
%   \caption{Equivalent (von Mises) stress distribution in the elliptic plate}
%   \label{fig:elliptic_plate_sigmaEq}
%\end{figure}
%
%

%\paragraph{Case 2: Narrow T-section component under tension}
%This case, proposed by \citet{Demirdzic1997a}, consists of a narrow engineering component with a T cross-section (Figure \ref{fig:narrowTmember}).
%The case is 3-D, static, with linear elastic material behaviour.
%Symmetry allows one-quarter of the geometry to be simulated.
%A constant negative pressure of 1 MPa is applied to the lower surface, and the upper left surface is fully clamped.
%The Young’s modulus is $E = 210$ GPa, and Poisson’s ratio is $\nu = 0.3$.
%A radius $5$ mm hole is located at the expected stress concentration.
%%\begin{figure}[htbp]
%%   \centering
%%   \includegraphics[width=0.4\textwidth]{figures/narrowTmember-geometry.pdf} 
%%   \caption{Inflation of an idealised ventricle case geometry, mesh and loading conditions}
%%   \label{fig:narrowTmember}
%%\end{figure}
%\begin{figure}[htbp]
%	\centering
%	\subfigure[Case geometry (considered quarter geometry)]
%	{
%		\label{fig:narrowTmember_geometry}
%   		\includegraphics[height=0.4\textwidth]{figures/narrowTmember-geometry} 
%   	}
%   	\qquad
%	\subfigure[Mesh with 624 cells]
%	{
%		\label{fig:narrowTmember_mesh}
%   		\includegraphics[height=0.4\textwidth]{figures/narrowTmember-mesh}  
%   	}
%	\caption{Narrow T-section component under tension: geometry and mesh}
%	\label{fig:narrowTmember}
%\end{figure}
%
%Five succesively refined hexahedral meshes are created using the OpenFOAM \texttt{blockMesh} utility, consisting of 624, $4\,992$, $39\,936$, $319\,488$, and $2\,555\,904$ cells.
%The corresponding average cell widths—calculated as the cubed root of the total domain volume divided by the number of cells—are 6.69, 3.34, 1.67, 0.84, and 0.42 mm.
%
%
%The predicted equivalent (von Mises) stress distribution on the $z = 0$ plane is shown in Figure \ref{fig:narrowTmember_sigmaEq}(a), where a stress concentration is apparent near the hole.
%The equivalent stress along the line $r = 1.5R$, $z = 0$ is shown for the different meshes in Figure \ref{fig:narrowTmember_sigmaEq}(b), where $R$ is the hole radius.
%%Similarly to the elliptic plate case, 
%The stress values have been sampled at 30 uniformly spaced angles between $-\frac{\pi}{2}$ and $-\pi$ along the line $r = 1.5R$, $z = 0$, finding the cell of each sampled point, and extrapolating from the cell-centred value using the field gradient.
%The predictions are seen converge to the results reported in \citet{Demirdzic1997a}.
%\begin{figure}[htbp]
%   \centering
%	\subfigure[Equivalent stress distribution for the mesh with $2\,555\,904$ cells on the $z = 0$ m plane]
%	{
%		\includegraphics[height=0.32\textwidth]{figures/narrowTmember_sigmaEq}  
%	}
%	\subfigure[Equivalent stress along the line $r = 1.5R$, $z = 0$ m]
%	{
%		\includegraphics[height=0.32\textwidth]{figures/narrowTmember_sigmaEq_line.pdf} 
%	}
%   \caption{Equivalent (von Mises) stress distribution in the narrow T-section component under tension}
%   \label{fig:narrowTmember_sigmaEq}
%\end{figure}

\paragraph{Case 3: Inflation of an idealised ventricle}
Inflation of an idealised ventricle (Figure \ref{fig:ventricle}) was proposed by \citet{Land2015} as a benchmark problem for cardiac mechanics software.
The case is 3-D, static, with finite hyperelastic strains.
The initial geometry is defined as a truncated ellipsoid:
\begin{eqnarray}
	x = r_s \sin(u) \cos(v), \quad
	y = r_s \sin(u) \sin(v), \quad
	z = r_l \cos(u)
\end{eqnarray}
where on the inner (endocardial) surface $r_s =7$ mm, $r_l = 17$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{17} \right) \right]$ and $v \in \left[-\pi, \pi \right]$, while on the outer (epicardial) surface $r_s =10$ mm, $r_l = 20$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{20} \right) \right]$ and $v \in \left[-\pi, \pi \right]$.
The base plane $z = 5$ mm is implicitly defined by the ranges for $u$.
%Constitutive parameters: isotropic, 
The hyperelastic material behaviour is described by the transversely isotropic constitutive law proposed by \citet{Guccione1995} law, where the parameters are $C = 10$ kPa, $c_f = c_t = c_{fs} = 1$; the specified parameters produce isotropic behaviour.
The benchmark specifies an incompressible material; in the current work, incompressibility is enforced weakly using a penalty approach, where the bulk modulus parameter is chosen to be two orders of magnitude greater ($\kappa = 1000$ kPa) than the greatest shear modulus parameter.
A pressure of 10 kPa is applied to the inner surface, applied here over 100 equal load increments, and the base plane is fixed.
The geometry is meshed using a structured approach and is predominantly composed of hexahedra, with prism cells forming the apex.
The OpenFOAM utilities \texttt{blockMesh} and \texttt{extrudeMesh} are used to create the meshes.
Four successively refined meshes are examined: $1\,620$ (shown in Figure \ref{fig:ventricle}(a)), $12\,960$, $103\,680$, and $829\,440$ cells.
The case can be simulated as 2-D axisymmetric but is simulated here using the full 3-D geometry.
%Figure \ref{fig:ventricle} shows the mesh with 1,620 cells (1,512 hexahedra and 108 prisms).
\begin{figure}[htbp]
   \centering
%   \includegraphics[width=0.4\textwidth]{figures/ventricle} 
	\subfigure[Ventricle undeformed geometry, showing the mesh with 1,620 cells]
	{
		\includegraphics[height=0.46\textwidth]{figures/ventricle} 
	}
	\subfigure[Cross-section showing the equivalent (von Mises) stress distribution for the mesh with 829,440 cells]
	{
		\includegraphics[height=0.48\textwidth]{figures/ventricle_sigmaEq}
	}
   \caption{Idealised ventricle case}
   \label{fig:ventricle}
\end{figure}

@@

\paragraph{Case 4: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
The 2-D plane strain tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
%\begin{figure}[htbp]
%   \centering
%   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry} 
%   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
%   \label{fig:cooks_membrane}
%\end{figure}
\begin{figure}[htbp]
	\centering
	\subfigure[Case geometry and dimensions (in mm)]
	{
		\label{fig:cooks_membrane_geometry}
   		\includegraphics[scale=1]{figures/cooksMembrane-geometry} 
   	}
   	\qquad
	\subfigure[Quadrilateral mesh with 144 cells]
	{
		\label{fig:cooks_membrane_mesh}
   		\includegraphics[scale=0.14]{figures/cooksMembrane-mesh}  
   	}
	\caption{Cook's membrane case geometry and mesh}
   \label{fig:cooks_membrane}
\end{figure}

@@

\paragraph{Case 5: Vibration of a 3-D Cantilevered Beam}
This 3-D, dynamic, finite strain case geometry consists of a $2 \times 0.2 \times 0.2$ m cuboid column and was proposed in its initial form by \citet{Tukovic2007}.
%(Figure \ref{fig:dynamic_cantilever})
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%\begin{eqnarray}
%	|\bb{T}| = \frac{\mathcal{F} E I}{A L^2} \text{ Pa}
%\end{eqnarray}
A sudden, constant traction $\bb{T} = \left(80, 80, 0 \right)$ kPa is applied to the upper surface.
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%In the current study, $\mathcal{F} = 1$.
A neo-Hookean hyperelastic material is assumed with $E = 15.293$ MPa, $\nu = 0.3$ and density $\rho = 1000$ kg m$^{-3}$, while the area $A = 0.04$ m$^2$ and the second moment of area $I = \nicefrac{1}{7\,500}$ m$^4$. %0.0001333
The geometric and material parameters were chosen for the first natural frequency to be 1 Hz in the small strain limit.
The magnitude of the applied traction is chosen here to test the robustness of the solution procedures when large rotations and strains are present.
Four succesively refined hexahedral meshes are generated using the OpenFOAM \texttt{blockMesh} utility, with cell counts of 270, $2\,160$, $17\,280$ and $138\,240$.
%A fixed time step size of 0.001617274174 s is chosen, corresponding to a Courant number of six on the mesh with $2\,160$ cells.
%A fixed time step size of 1 ms is chosen.
The time step size is 1 ms, and the total period is 1 s.
% corresponding to 10 expected oscillation periods.
%, corresponding to a Courant number of six on the mesh with $2\,160$ cells.

>>>>

It is noted that the displacement is assumed to vary linearly within each cell; hence, the displacement gradient and stress are constant within each cell.
In the current work, a unique definition of stress $\bb{\sigma}_{f_i}$ at each cell face $f_i$ is given as a weighted-averaged of values in the two cells ($\bb{\sigma}_P $, $\bb{\sigma}_{N_{f_i}}$) straddling the face \citep{Jasak1996}:
%The stress $\bb{\sigma}_{f_i}$ at an internal face is calculated by linear interpolation from the adjacent cell centres \citep{Jasak1996}:
\begin{eqnarray} \label{eq:stressInterp}
	\bb{\sigma}_{f_i} &=& w_{f_i} \bb{\sigma}_P + (1 - w_{f_i}) \bb{\sigma}_{N_{f_i}}
\end{eqnarray}
%where $\bb{\sigma}_{f_i}$ is the stress at an interface face $f_i$, 
where the interpolation weight is defined as $w_{f_i} = (\bb{n}_{f_i} \cdot [\bb{x}_{N_{f_i}} - \bb{x}_{f_i}])/(\bb{n}_{f_i} \cdot [\bb{x}_{N_{f_i}} - \bb{x}_{P} ])$; however, achieving second-order accuracy of the displacement field is independent of the value of the weights, and, for example, $w_{f_i} = \nicefrac{1}{2}$ would also be sufficient.

The stress $\bb{\sigma}_{s_i}$ at a symmetry boundary face is calculated as
\begin{eqnarray} \label{eq:symm}
	\bb{\sigma}_{s_i}
		&=& \frac{1}{2} \left (\bb{\sigma}_P + \bb{R}_{s_i} \cdot \bb{\sigma}_{P} \right) \notag \\
		&=& \left (\textbf{I} - \bb{n}_{s_i} \otimes \bb{n}_{s_i} \right) \cdot \bb{\sigma}_P
\end{eqnarray}
where $\bb{R}_{s_i} \cdot \bb{\sigma}_{P}$ represents the mirror reflection of $\bb{\sigma}_P$ across the symmetry boundary face $s_i$.
The reflection tensor is $\bb{R}_{s_i} = \textbf{I} - 2 \bb{n}_{s_i} \otimes \bb{n}_{s_i}$ \citep{Demirdzic2022}, with $\bb{n}_{s_i}$ indicating the unit normal of the symmetry boundary face $s_i$.
From Equation \ref{eq:symm}, it is clear that shear stresses are zero on a symmetry plane boundary face.
Note from Equation \ref{eq:divStressDiscret} that the stress on displacement boundary faces is assumed to be equal to the stress $\bb{\sigma}_P$ at the centroid of cell $P$.

@@

\hl{magSf in OF/ d3 weights}

@@

first-order accurate gradients are sufficient to preserve second-order accuracy of the cell-centre displacements. \hl{OF uses magSf}

@@

\paragraph{Case 3: Out-of-plane bending of an elliptic plate}
This 3-D, static, linear elastic test case (Figure \ref{fig:elliptic_plate}) consists of a thick elliptic plate (0.6 m thick) with a centred elliptic hole, with the inner and outer ellipses given as
\begin{eqnarray}
	\left(\frac{x}{2}\right)^2 + \left(\frac{y}{1}\right)^2 = 1 & \text{inner ellipse} \\
	\left(\frac{x}{3.25}\right)^2 + \left(\frac{y}{2.75}\right)^2 = 1 & \text{outer ellipse}
\end{eqnarray}
The case has been described by the National Agency for Finite Element Methods and Standards (NAFEMS) \cite{Hitchings1987}, and analysed using finite volume procedures by \citet{Demirdzic1997a} and \citet{Cardiff2016a}.
Symmetry allows one-quarter of the geometry to be simulated.
A constant pressure of 1 MPa is applied to the upper surface, and the outer surface is fully clamped.
The mechanical properties are: $E = 210$ GPa, $\nu = 0.3$.
Six successively refined hexahedral meshes are used, with cell counts of 45, 472 (Figure \ref{fig:elliptic_plate}), $4\,140$, $34\,968$, $287\,280$ and $2\,438\,242$.
\begin{figure}[htbp]
   \centering
%\includegraphics[width=0.5\textwidth]{figures/elliptic_plate.pdf} ellipticPlate-geometry
%	\subfigure[Mesh containing 472 hexahedral cells]
%	{
		\includegraphics[width=0.45\textwidth]{figures/elliptic_plate.pdf} 
%	}
   \caption{Elliptic Plate geometry and mesh containing 472 hexahedral cells}
   \label{fig:elliptic_plate}
\end{figure}

The prediction for the equivalent (von Mises) stress in the domain are shown in Figure \ref{fig:elliptic_plate_sigmaEq}(a), and the values along the line $r = \sqrt{x^2 + y^2} = 2.1$ m, $z = 0.3$ m are shown in Figure \ref{fig:elliptic_plate_sigmaEq}(b).
The results from the finest grid in \citet{Demirdzic1997a} are given for comparison, where good agreement is seen;
the small offset between the finest mesh predictions and those from \citet{Demirdzic1997a} are likely due to errors introduced when extracting the \citet{Demirdzic1997a} results using WebPlotDigitizer \citep{WebPlotDigitizer}.
The sample values have been calculated by generating 40 uniformly spaced angles between 0 and $\nicefrac{\pi}{2}$ along the line, finding the cell of each sampled point, and extrapolating from the cell-centred value using the field gradient.
\begin{figure}[htbp]
   \centering
   %\includegraphics[width=0.5\textwidth]{figures/elliptic_plate_sigmaEq_line.pdf} 
	\subfigure[Equivalent stress distribution for the mesh with $2\,438\,242$ cells. The deformation is scaled by $1\,000$ and a translucent line indicates the outline of the undeformed geometry]
	{
		\includegraphics[width=0.45\textwidth]{figures/elliptic_plate_sigmaEq}  
	}
	\subfigure[Equivalent stress along the line $r = 2.1$, $z = 0.3$ m]
	{
		\includegraphics[width=0.45\textwidth]{figures/elliptic_plate_sigmaEq_line.pdf} 
	}
   \caption{Equivalent (von Mises) stress distribution in the elliptic plate}
   \label{fig:elliptic_plate_sigmaEq}
\end{figure}

%\paragraph{Case X: Narrow T-section component under tension}
%This case, proposed by \citet{Demirdzic1997a}, consists of a narrow engineering component with a T cross-section (Figure \ref{fig:narrowTmember}).
%The case is 3-D, static, with linear elastic material behaviour.
%Symmetry allows one-quarter of the geometry to be simulated.
%A constant negative pressure of 1 MPa is applied to the lower surface, and the upper left surface is fully clamped.
%The Young’s modulus is $E = 210$ GPa, and Poisson’s ratio is $\nu = 0.3$.
%A radius $5$ mm hole is located at the expected stress concentration.
%%\begin{figure}[htbp]
%%   \centering
%%   \includegraphics[width=0.4\textwidth]{figures/narrowTmember-geometry.pdf} 
%%   \caption{Inflation of an idealised ventricle case geometry, mesh and loading conditions}
%%   \label{fig:narrowTmember}
%%\end{figure}
%\begin{figure}[htbp]
%	\centering
%	\subfigure[Case geometry (considered quarter geometry)]
%	{
%		\label{fig:narrowTmember_geometry}
%   		\includegraphics[height=0.4\textwidth]{figures/narrowTmember-geometry} 
%   	}
%   	\qquad
%	\subfigure[Mesh with 624 cells]
%	{
%		\label{fig:narrowTmember_mesh}
%   		\includegraphics[height=0.4\textwidth]{figures/narrowTmember-mesh}  
%   	}
%	\caption{Narrow T-section component under tension: geometry and mesh}
%	\label{fig:narrowTmember}
%\end{figure}
%
%Five succesively refined hexahedral meshes are created using the OpenFOAM \texttt{blockMesh} utility, consisting of 624, $4\,992$, $39\,936$, $319\,488$, and $2\,555\,904$ cells.
%The corresponding average cell widths—calculated as the cubed root of the total domain volume divided by the number of cells—are 6.69, 3.34, 1.67, 0.84, and 0.42 mm.
%
%
%The predicted equivalent (von Mises) stress distribution on the $z = 0$ plane is shown in Figure \ref{fig:narrowTmember_sigmaEq}(a), where a stress concentration is apparent near the hole.
%The equivalent stress along the line $r = 1.5R$, $z = 0$ is shown for the different meshes in Figure \ref{fig:narrowTmember_sigmaEq}(b), where $R$ is the hole radius.
%%Similarly to the elliptic plate case, 
%The stress values have been sampled at 30 uniformly spaced angles between $-\frac{\pi}{2}$ and $-\pi$ along the line $r = 1.5R$, $z = 0$, finding the cell of each sampled point, and extrapolating from the cell-centred value using the field gradient.
%The predictions are seen converge to the results reported in \citet{Demirdzic1997a}.
%\begin{figure}[htbp]
%   \centering
%	\subfigure[Equivalent stress distribution for the mesh with $2\,555\,904$ cells on the $z = 0$ m plane]
%	{
%		\includegraphics[height=0.32\textwidth]{figures/narrowTmember_sigmaEq}  
%	}
%	\subfigure[Equivalent stress along the line $r = 1.5R$, $z = 0$ m]
%	{
%		\includegraphics[height=0.32\textwidth]{figures/narrowTmember_sigmaEq_line.pdf} 
%	}
%   \caption{Equivalent (von Mises) stress distribution in the narrow T-section component under tension}
%   \label{fig:narrowTmember_sigmaEq}
%\end{figure}

\paragraph{Case 4: Inflation of an idealised ventricle}
Inflation of an idealised ventricle (Figure \ref{fig:ventricle}) was proposed by \citet{Land2015} as a benchmark problem for cardiac mechanics software.
The case is 3-D, static, with finite hyperelastic strains.
The initial geometry is defined as a truncated ellipsoid:
\begin{eqnarray}
	x = r_s \sin(u) \cos(v), \quad
	y = r_s \sin(u) \sin(v), \quad
	z = r_l \cos(u)
\end{eqnarray}
where on the inner (endocardial) surface $r_s =7$ mm, $r_l = 17$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{17} \right) \right]$ and $v \in \left[-\pi, \pi \right]$, while on the outer (epicardial) surface $r_s =10$ mm, $r_l = 20$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{20} \right) \right]$ and $v \in \left[-\pi, \pi \right]$.
The base plane $z = 5$ mm is implicitly defined by the ranges for $u$.
%Constitutive parameters: isotropic, 
The hyperelastic material behaviour is described by the transversely isotropic constitutive law proposed by \citet{Guccione1995} law, where the parameters are $C = 10$ kPa, $c_f = c_t = c_{fs} = 1$; the specified parameters produce isotropic behaviour.
The benchmark specifies an incompressible material; in the current work, incompressibility is enforced weakly using a penalty approach, where the bulk modulus parameter is chosen to be two orders of magnitude greater ($\kappa = 1000$ kPa) than the greatest shear modulus parameter.
A pressure of 10 kPa is applied to the inner surface, applied here over 100 equal load increments, and the base plane is fixed.
The geometry is meshed using a structured approach and is predominantly composed of hexahedra, with prism cells forming the apex.
The OpenFOAM utilities \texttt{blockMesh} and \texttt{extrudeMesh} are used to create the meshes.
Four successively refined meshes are examined: $1\,620$ (shown in Figure \ref{fig:ventricle}(a)), $12\,960$, $103\,680$, and $829\,440$ cells.
The case can be simulated as 2-D axisymmetric but is simulated here using the full 3-D geometry.
%Figure \ref{fig:ventricle} shows the mesh with 1,620 cells (1,512 hexahedra and 108 prisms).
\begin{figure}[htbp]
   \centering
%   \includegraphics[width=0.4\textwidth]{figures/ventricle} 
	\subfigure[Ventricle undeformed geometry, showing the mesh with 1,620 cells]
	{
		\includegraphics[height=0.46\textwidth]{figures/ventricle} 
	}
	\subfigure[Cross-section showing the equivalent (von Mises) stress distribution for the mesh with 829,440 cells]
	{
		\includegraphics[height=0.48\textwidth]{figures/ventricle_sigmaEq}
	}
   \caption{Idealised ventricle case}
   \label{fig:ventricle}
\end{figure}

@@

\paragraph{Case 5: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
The 2-D plane strain tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
%\begin{figure}[htbp]
%   \centering
%   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry} 
%   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
%   \label{fig:cooks_membrane}
%\end{figure}
\begin{figure}[htbp]
	\centering
	\subfigure[Case geometry and dimensions (in mm)]
	{
		\label{fig:cooks_membrane_geometry}
   		\includegraphics[scale=1]{figures/cooksMembrane-geometry} 
   	}
   	\qquad
	\subfigure[Quadrilateral mesh with 144 cells]
	{
		\label{fig:cooks_membrane_mesh}
   		\includegraphics[scale=0.14]{figures/cooksMembrane-mesh}  
   	}
	\caption{Cook's membrane case geometry and mesh}
   \label{fig:cooks_membrane}
\end{figure}

@@

\paragraph{Case 6: Vibration of a 3-D Cantilevered Beam}
This 3-D, dynamic, finite strain case geometry consists of a $2 \times 0.2 \times 0.2$ m cuboid column and was proposed in its initial form by \citet{Tukovic2007}.
%(Figure \ref{fig:dynamic_cantilever})
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%\begin{eqnarray}
%	|\bb{T}| = \frac{\mathcal{F} E I}{A L^2} \text{ Pa}
%\end{eqnarray}
A sudden, constant traction $\bb{T} = \left(80, 80, 0 \right)$ kPa is applied to the upper surface.
%A sudden, constant traction $\bb{T} = \left(\sfrac{|\bb{T}|}{\sqrt{2}}\right) \left(1, 1, 0 \right)$ Pa is applied to the upper surface, where the magnitude $|\bb{T}|$ is defined in terms of a dimensionless load factor $\mathcal{F}$ as
%In the current study, $\mathcal{F} = 1$.
A neo-Hookean hyperelastic material is assumed with $E = 15.293$ MPa, $\nu = 0.3$ and density $\rho = 1000$ kg m$^{-3}$, while the area $A = 0.04$ m$^2$ and the second moment of area $I = \nicefrac{1}{7\,500}$ m$^4$. %0.0001333
The geometric and material parameters were chosen for the first natural frequency to be 1 Hz in the small strain limit.
The magnitude of the applied traction is chosen here to test the robustness of the solution procedures when large rotations and strains are present.
Four succesively refined hexahedral meshes are generated using the OpenFOAM \texttt{blockMesh} utility, with cell counts of 270, $2\,160$, $17\,280$ and $138\,240$.
%A fixed time step size of 0.001617274174 s is chosen, corresponding to a Courant number of six on the mesh with $2\,160$ cells.
%A fixed time step size of 1 ms is chosen.
The time step size is 1 ms, and the total period is 1 s.
% corresponding to 10 expected oscillation periods.
%, corresponding to a Courant number of six on the mesh with $2\,160$ cells.
